### 2018-09-06 tex/article.tex
add paper draft and docs

\documentclass[twocolumn]{svjour3}
\usepackage{style}
\usepackage{subfiles}
\usepackage{microtype}

% Title
\title{Multiset-trie data structure}
\author{Mikita Akulich \and Iztok Savnik \and Matja\v z Krnc \and Riste \v Skrekovski}
\institute{
		Mikita Akulich 
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{mikita.akulich@gmail.com}
		\and
		Iztok Savnik
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{iztok.savnik@famnit.upr.si}
		\and 
		Matja\v z Krnc 
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{matjaz.krnc@famnit.upr.si}
		\and 
		Riste \v Skrekovski
		\at
		University of Ljubljana, FMF, Ljubljana, Slovenia\\
		\email{riste.skrekovski@famnit.upr.si}
		}

\begin{document}
\maketitle

%Abstract
\begin{abstract}
[Will be rewritten]
In this paper we will present the multiset-trie, a new data structure that operates 
on objects represented as multisets. The multiset-trie is a search-tree-based 
data structure with the properties similar to those of a trie. It implements all standard search 
tree operations together with the special multiset containment operations. Multiset 
containment operations supported by the multiset-trie are submultiset and supermultiset. 
These operations are used for implementation of different queries that can be performed 
on multisets in a multiset-trie. One of the most important queries is the search of 
the nearest neighbor given an input object. The nearest neighbor search of a 
multiset-trie makes it a good alternative for the index data structures that are 
used in information retrieval systems. In particular, our research is focused on the 
application of the multiset-trie to full-text search systems. 
\end{abstract}
\keywords{multiset, bag-of-words, trie, multiset-trie, information retrieval, inverted index, full-text search}
%\subclass{68P05, 68P20, 68Q87}

% section 1
\input{tex/s_intro}
% section 2
\input{tex/s_data_structure}
% section 3
\input{tex/s_operations}
% section 4
\input{tex/s_math}
% section 5
\input{tex/s_experiments}
% section 6
\input{tex/s_relwork}
% section 7
\input{tex/s_conclusions}

\bibliographystyle{abbrv}
\bibliography{multiset-trie}

\end{document}