### 2018-09-06 tex/tex/s_operations.tex

\section{Multiset-trie operations} \label{c:operations}
%
Let $\mathcal{M}$ be a multiset-trie and let $M$ be a set of multisets that are 
inserted into the multiset-trie $\mathcal{M}.$ We define a type \emph{Multiset} in 
order to use it as a representation of a multiset. The type \emph{Multiset} is 
an array $m$ of constant length $\sigma,$ where $i-th$ cell represents the element 
$\phi^{-1}(i)$ from $\Sigma$ with multiplicity $m[i].$ From now on we 
agree that the first cell of an array has index 1. Let us have an example of a 
\emph{Multiset} instance with $\sigma = 2:$
%
\begin{center}
\begin{tabular}{ccc}
Multiset & & Instance of type Multiset \\
$\{ 1,1,2 \}$ & $\cong $ & 
\begin{tabular}{|c|c|}
\hline 
2 & 1 \\
\hline 
\multicolumn{1}{c}{\tiny 1} & \multicolumn{1}{c}{\tiny 2} \\
\end{tabular}
\end{tabular}
\end{center}
%
The operations supported by the multiset-trie data structure are as follows.
%
\begin{enumerate}
\item \textsc{insert}($\mathcal{M}$, $m$): inserts a multiset $m$ into 
$\mathcal{M}$ if $m\not\in M;$
%
\item \textsc{search}($\mathcal{M}$, $m$): returns true if a multiset $m\in M$ 
for a given $\mathcal{M},$ and returns false otherwise;
%
\item \textsc{delete}($\mathcal{M}$, $m$): returns true if a multiset $m$ was 
successfully deleted from $\mathcal{M},$ and returns false otherwise (in case 
$m\not\in M$);
%
\item \textsc{submsetExistence}($\mathcal{M}$, $m$): returns true if 
there exists a $x\in M$ for a given $\mathcal{M}$ such that $x\subseteq m,$ 
and returns false otherwise; 
%
\item \textsc{supermsetExistence}($\mathcal{M}$, $m$): returns true if 
there exists a $x\in M$ for a given $\mathcal{M}$ such that $x\supseteq m,$ and returns 
false otherwise; 
%
\item \textsc{getAllSubmsets}($\mathcal{M}$, $m$): returns the set of multisets 
$\{ x \in M : x\subseteq m \}$ for a given $\mathcal{M};$
%
\item \textsc{getAllSupermsets}($\mathcal{M}$, $m$): returns the set of multisets 
$\{ x\in M : x\supseteq m \}$ for a given $\mathcal{M}.$
%
\end{enumerate}

In the following subsections we will present each operation of the multiset-trie 
data structure separately.

Firstly we would like to describe some notations that will be used. The 
multiset-trie data structure is a recursive data structure. Hence, any sub 
tree of a multiset-trie $\mathcal{M}$ is again a multiset-trie. This fact allows 
us to use the root node of a multiset-trie as its representative. Thus, the notation 
$\mathcal{M}$ will be used instead of $\mathcal{M}.root$ to refer to the root 
node of $\mathcal{M}.$ Non-existing or \emph{Null} nodes in multiset-trie will 
be marked as \emph{Null} and existing nodes at the level $LL$ will be marked 
as \emph{accepting} nodes. The array slicing operation will be used as follows. 
For a given array $a,$ $a[i:]$ represents the array obtained from $a$ by taking 
only the cells from index $i$ until the last cell.

\subsection{Insert} \label{s:insert}
The procedure \textsc{insert}($\mathcal{M}$, $m$) inserts a new instance $m$ of 
type Multiset into multiset-trie $\mathcal{M}$. If the complete path already 
exists, then procedure leaves the structure unchanged. Otherwise it extends 
partially existing or creates a new complete path. The procedure does not return 
any result. The pseudocode for procedure \textsc{insert} is presented in 
Algorithm~\ref{alg:insert}.

\begin{algorithm}[h!]
\caption{Procedure \textsc{insert}}
\label{alg:insert}
\begin{algorithmic}[1]
\Procedure{\textsc{insert}}{$\mathcal{M}$, $m$}
\State $currentNode \gets \mathcal{M}$
\For{$i=1$ to $\sigma$}
\If {child $c_{m[i]}$ of $currentNode$ is \emph{Null}}
\State create new child $c_{m[i]}$ of $currentNode$
\EndIf
\State $currentNode \gets c_{m[i]}$
\EndFor
\State mark $currentNode$ as \emph{accepting}
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Search}\label{s:search}
The function \textsc{search}($\mathcal{M}$, $m$) checks if the complete path corresponding to 
a given multiset $m$ exists in the structure $\mathcal{M}.$ The function returns 
true if the multiset $m$ exists in $\mathcal{M}$, and returns false otherwise. The 
function \textsc{search} is presented in Algorithm~\ref{alg:search}.

\begin{algorithm}[h!]
\caption{Function \textsc{search}}
\label{alg:search}
\begin{algorithmic}[1]
\Function{\textsc{search}}{$\mathcal{M}$, $m$}
\State $currentNode \gets \mathcal{M}$
\For{$i=1$ to $\sigma$}
\If {child $c_{m[i]}$ of $currentNode$ is \emph{Null}}
\State \Return False
\EndIf
\State $currentNode \gets c_{m[i]}$
\EndFor
\State \Return True
\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Delete} \label{s:delete}
The function \textsc{delete}($\mathcal{M},$ $m$) searches for the complete path 
that corresponds to $m$ in order to remove it. If the path can not be found, the 
function immediately returns false. During search, the function keeps track of the 
number of children for every node. It marks the nodes that have more than one child 
as \emph{parent nodes} and remembers the label of the child which is a potential node 
where the sub-tree will be cut to remove the multiset. The parent node is needed to 
perform a removal, because the multiset-tire is an explicit data structure. When search 
is completed, the function removes the sub-tree of the last found parent node, and 
returns true. In such a way after deletion all the prefixes for other multisets are 
preserved in $\mathcal{M}$ and $m$ is removed. The function \textsc{delete} is 
presented in Algorithm~\ref{alg:delete}.

\begin{algorithm}[h!]
\caption{Function \textsc{delete}}
\label{alg:delete}
\begin{algorithmic}[1]
\Function{\textsc{delete}}{$\mathcal{M},$ $m$}
\State $currentNode \gets \mathcal{M}$
\State $parent \gets currentNode$ 
\State $position \gets 1$
\For {$i=1$ to $\sigma$}
\If {child $c_{m[i]}$ of $currentNode$ is \emph{Null}}
\State \Return False
\EndIf
\State $numChildren \gets 0$
\For {$j=0$ to $n-1$}
\If {child $c_j$ of $currentNode$ is not \emph{Null}}
\State $numChildren\gets numChildren+1$
\EndIf
\EndFor
\If {$numChildren$ is not $1$}
\State $parent\gets currentNode$
\State $position \gets i$
\EndIf
\State $currentNode \gets c_{m[i]}$
\EndFor
\State child $c_{m[position]}$ of $parent\gets$ \emph{Null}
\State \Return True
\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Sub-multiset existence} \label{s:subexists}
The function \textsc{submsetExistence}($\mathcal{M}$, $m$) checks if there exists 
a multiset $x$ in $\mathcal{M},$ that satisfies the condition $x\subseteq m.$ 
The function starts with searching for an exact match $x=m$ in $\mathcal{M},$ 
since $m\subseteq m$ by definition of submultiset inclusion. If an exact match is 
not found in $\mathcal{M},$ the function uses multiset-trie to find the closest 
(the largest) submultiset of $m$ in $\mathcal{M}$ by decreasing multiplicity of 
elements in $m.$ At every level the function tries to proceed with the largest 
possible multiplicity of an element that is provided by $m.$ However, when the 
function reaches some level where it meets a \emph{Null} node and can not go 
further using path provided by $m,$ it decreases the multiplicity of an element 
that corresponds to a current level. Thus, the function can decrease multiplicity 
of an element or eventually skip it in order to find the closest $x\subseteq m.$ 
The function \textsc{submsetExistence} is presented in Algorithm~\ref{alg:subexists}.

\begin{algorithm}[h!]
\caption{Function \textsc{submsetExistence}}
\label{alg:subexists}
\begin{algorithmic}[1]
\Function{\textsc{submsetExistence}}{$\mathcal{M}$, $m$}
\State $currentNode \gets \mathcal{M}$
\If {$currentNode$ is \emph{accepting}}
\State \Return True
\EndIf
\For {$i=m[1]$ down to $0$}
\If {child $c_i$ of $currentNode$ is not \emph{Null}}
\If {\textsc{submsetExistence}($c_i,$ $m[2:]$)}
\State \Return True
\EndIf
\EndIf
\EndFor
\State \Return False
\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Super-multiset existence} \label{s:superexists}
The function \textsc{supermsetExistence}($\mathcal{M},$ $m$) checks if there 
exists supermultiset $x$ of a given multiset $m$ in $\mathcal{M}.$ By analogy 
to the function \textsc{submsetExistence}, the function \textsc{supermsetExistence} 
starts by searching for an exact match $x=m$ in $\mathcal{M}.$ If an exact 
match is not found in $\mathcal{M},$ the function searches for the closest (the 
smallest) supermultiset $x$ of $m$ in $\mathcal{M}$ by increasing multiplicity 
of elements in $m.$ At every level the function tries to proceed with the smallest 
possible multiplicity of an element that is provided by $m.$ However, when function 
reaches some level where it meets a \emph{Null} node and can not go further 
using path provided by $m,$ it increases the multiplicity of an element that 
corresponds to a current level. Thus, the function \textsc{supermsetExistence} 
can increase multiplicity of an element up to $n-1,$ where $n$ is the degree of 
a node in $\mathcal{M},$ to find the closest supermultiset $x\supseteq m$ in 
$\mathcal{M}.$ The function \textsc{supermsetExistence} is presented in 
Algorithm~\ref{alg:superxists}.

\begin{algorithm}[h!]
\caption{Function \textsc{supermsetExistence}}
\label{alg:superxists}
\begin{algorithmic}[1]
\Function{\textsc{supermsetExistence}}{$\mathcal{M}$, $m$}
\State $currentNode \gets \mathcal{M}$
\If {$currentNode$ is \emph{accepting} }
\State \Return True
\EndIf
\For {$i=m[1]$ to $n-1$}
\If {child $c_i$ of $currentNode$ is not \emph{Null}}
\If {\textsc{supermsetExistence}($c_i,$ $m[2:]$)}
\State \Return True
\EndIf
\EndIf
\EndFor
\State \Return False
\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Get all sub-multisets and get all super-multisets} \label{s:getall}
The algorithms for functions \textsc{getAllSubmsets} and 
\textsc{getAllSupermsets} are based entirely on algorithms for 
\textsc{submsetExistence} and \textsc{supermsetExistence} functions that do not 
terminate on the first existing sub/supermultiset, but store the results and 
continue procedure until all existing sub/supermultisets in $\mathcal{M}$ are 
found and stored.