### 2018-09-06 tex/tex/s_relwork.tex

\section{Related work} \label{c:relwork}
[Will be rewritten]
%
% Set theory, multiest theory, bag-of-words model
\subsection{Multiset}
Multiset is a widely used data structure in different areas of mathematics, physics 
and computer science \cite{singh2007overview}. The theory of multisets is based 
entirely on the theory of sets. However, classical mathematics does not deal with 
multisets directly. Instead, one can define a multiset to be a \emph{family} of sets 
or the functions on ordered pairs, where the members of a pair are an element and its 
multiplicity. This means that mathematically the concepts of set such as cardinality, 
set-containment operation, power set, equivalence classes and others are well defined 
for multisets in terms of sets \cite{blizard1988multiset}.

The concept of a multiset can also be referred to the \emph{bag-of-words model}. 
This model takes its origin from a linguistic context studied by Harris~\cite{harris1954distributional}. 
According to the bag-of-words model, text can be represented as a bag (multiset) 
of words, where an element is a word and the number of its occurrences in the 
text is multiplicity. A bag of words does not keep track of grammar and ordering 
of words.

% Information retrieval
\subsection{Information retrieval}
\emph{Information retrieval (IR)} refers to a problem of finding material of an 
unstructured nature that satisfies an information need~\cite{manning2008introduction}. 
Usually, one is searching for a specific documents in a significantly large text 
documents database. The size of a database makes the search a time consuming 
operation. In order to resolve the issue IR systems pre-process data and create 
indexes for future use in search operation.

The bag-of-words model is widely used in IR. In particular, such a representation of 
text documents is used in database indexes when a full text search of a database 
is required. The full-text search problem refers to indexing techniques for full-text 
databases. The most efficient index nowadays uses the concept of an inverted index 
\cite{zobel1992efficient}.

The proposed data structure multiset-trie can be used as an alternative implementation 
of the search structure of an inverted index. It represents words as multisets and 
stores them into data structure. The query processing is achieved using boolean 
retrieval model \cite{manning2008introduction} and multiset containment operations. 
Multiset containment operations of the multiset-trie implement the nearest neighbor 
search queries which retrieve not only exact but also the most relevant results 
to a user.

\subsection{Generalized search tree}
The properties and operations of the multiset-trie makes it a competitor to the 
most efficient implementation of a search tree the \emph{Generalized Search Tree 
(GiST)}~\cite{broder2006indexing,hellerstein1995generalized,kornacker1999high} 
that is used in inverted index. GiST is a very flexible data structure that can be 
customized in order to behave like B+-tree, R-tree or RD-tree. It also provides 
support for an extensible set of queries and data types that B+-tree, R-tree or RD-tree 
do not support originally. GiST supports all the basic search tree operations such as 
insert, delete and search, and in addition provides such extensions as the 
nearest-neighbor search and multiset containment operations. The extensions 
provided by GiST are native in the multiset-trie. The multiset-trie also has a fixed 
height while GiST is a self-balanced tree and has to use additional methods in 
order to preserve its balance.

% Set-trie
\subsection{Set-trie}
The multiset containment queries are well studied in the area of relational 
databases. The queries are well-defined in the context of relational 
algebra~\cite{Lamperti2000msetsinds}. This problem was previously studied 
for a restricted case of queries. In particular, the storage and fast retrieval of sets 
were previously accomplished in data structure \emph{set-trie} proposed by 
Savnik~\cite{savnik2013index}.

The set-trie data structure is based on a trie data structure. It supports set 
containment operations such as retrieval of the nearest sub- and supersets and 
retrieval of all sub- and supersets from the data structure.

The data structure multiset-trie adapts the properties of the set-trie implementing 
the functions \textsc{submsetExistence}, \textsc{supermsetExistence}, 
\textsc{getAllSubmsets} and \textsc{getAllSupermsets} together with the basic 
tree functions such as \textsc{insert}, \textsc{delete} and \textsc{search}. 
Moreover, multiset-trie extends the abilities of the set-trie allowing to store and 
retrieve multisets. The downside of such an extension is that multiset-trie no 
longer supports path compression that was obtained in set-trie. However, the 
design of multiset-trie provides a constant worst case time complexity of search 
function independently of user input.