### 2019-02-18 tex/article.tex
article changes

\documentclass[twocolumn]{svjour3}
\usepackage{style}
\usepackage{subfiles}
\usepackage{microtype}

% Title
\title{Multiset-trie data structure}
\author{Mikita Akulich \and Iztok Savnik \and Matja\v z Krnc \and Riste \v Skrekovski}
\institute{
		Mikita Akulich 
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{mikita.akulich@gmail.com}
		\and
		Iztok Savnik
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{iztok.savnik@famnit.upr.si}
		\and 
		Matja\v z Krnc 
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{matjaz.krnc@famnit.upr.si}
		\and 
		Riste \v Skrekovski
		\at
		University of Ljubljana, FMF, Ljubljana, Slovenia\\
		\email{riste.skrekovski@famnit.upr.si}
		}

\begin{document}
\maketitle

%Abstract
\begin{abstract}
This work proposes a new data structure \emph{multiset-trie} that is designed for storage 
and efficient processing of a set of multisets. Moreover, multiset-trie can operate on 
a set of sets without efficiency loss.
%
The multiset-trie is a search-tree-based data structure with the properties similar to those of 
a trie. It implements all standard search tree operations together with the multiset containment 
operations such as submultiset and supermultiset. Containment operations allow efficient search and 
retrieval of the individual multisets and also collections of multsets. In particular, it can find and retrieve 
closest submultisets or supermultisets as well as all submultisets or supermultisets that are parameterized
by user input.
%
Multiset-trie is compared to modern solutions that provide containment operations on multisets. 
The study shows how efficient multiset-trie is by outperforming inverted file in both equality and 
containment queries by up to 4 orders of magnitude in query execution time.
\end{abstract}
\keywords{multiset-trie, multiset, bag-of-words, trie, data structure, containment operations, inverted file}
%\subclass{68P05, 68P20, 68Q87}

% section 1
\input{tex/s_intro}
% section 2
\input{tex/s_data_structure}
% section 3
\input{tex/s_operations}
% section 4
\input{tex/s_math}
% section 5
\input{tex/s_experiments}
% section 6
\input{tex/s_relwork}
% section 7
\input{tex/s_conclusions}

\bibliographystyle{abbrv}
\bibliography{multiset-trie}

\end{document}