### 2021-09-10 tex/article_els.tex
elsevier style

%\documentclass[twocolumn,numbook]{svjour3}
\documentclass[preprint]{elsarticle}

%\usepackage{times}
%\normalfont
%\usepackage{latexsym}
%\usepackage[T1]{fontenc}
%\usepackage[noend]{algpseudocode}
%\usepackage{algorithm}
%\usepackage{verbatim}
%\usepackage{amsfonts}    % amssymb}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{stmaryrd}
%\usepackage{wrapfig}

\usepackage{style}
\usepackage{subfiles}
\usepackage{microtype}
\usepackage[capitalise]{cleveref}

%\usepackage{graphics}
%\usepackage{graphicx}

% for math enviroments
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proof}{Proof}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
%\newtheorem{ex}[thm]{Exercise}
%\newtheorem{obs}[thm]{Observation}
%\newtheorem{example}[thm]{Example}

% oštevilčenje. To be commented out in the end
\pagestyle{plain}

\begin{document}

\begin{frontmatter}
\title{Multiset-trie data structure}
%
%\titlerunning{}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%

\author[FAMNIT]{Mikita Akulich}
\ead{iztok.savnik@upr.si}

\author[FAMNIT,FIS]{Iztok Savnik}
\ead{iztok.savnik@upr.si}

\author[FAMNIT]{Matja\v z Krnc\corref{cor1}} 
\ead{matjaz.krnc@upr.si}

\author[FMF,FIS]{Riste \v Skrekovski}
\ead{skrekovski@gmail.com}

%
%\authorrunning{Iztok Savnik} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Iztok Savnik}
%
\address[FAMNIT]{Faculty of mathematics, natural sciences and information technologies, University of Primorska, Slovenia}

\address[FMF]{Department of Mathematics, University of Ljubljana, Slovenia}

\address[FIS]{Faculty of Information Studies, Novo Mesto, Slovenia}

\cortext[cor1]{Corresponding author}

%\maketitle              % typeset the title of the contribution

\begin{abstract}
This paper proposes a new data structure \emph{multiset-trie} that is designed for storing and efficiently processing of a set of multisets. Moreover, multiset-trie can operate on a set of sets without efficiency loss.
%
The multiset-trie is a search tree with properties similar to those of a trie. It implements all standard search tree operations together with the multiset containment operations such as sub-multiset and super-multiset. Suppose we have a set of multisets $S$ and a multiset $X$. The multiset containment operations retrieve multisets from $S$ that are either sub-multisets or super-multisets of $X$.

We present the mathematical analysis of a multiset-trie that gives the time complexity of the algorithms and the space complexity of the data structure. Further, the empirical analysis of the data structure is implemented in a series of experiments. The experiments illuminate the time complexity space of the multiset containment operations. Finally, the comparison of the multiset-trie with the inverted index shows that the multiset-trie outperforms the inverted index by up to 4 orders of magnitude.
\end{abstract}
% 
\begin{keyword}
multiset, index data structure, containment operations, inverted file
\end{keyword}

\end{frontmatter}

% section 1
\input{tex/s_intro}
% section 2
\input{tex/s_data_structure}
% section 3
\input{tex/s_operations}
% section 4
\input{tex/s_math}
% section 5
\input{tex/s_experiments}
% section 6
\input{tex/s_relwork}
% section 7
\input{tex/s_conclusions}

\bibliographystyle{abbrv}
\bibliography{multiset-trie}

\newpage
\tableofcontents

\end{document}