### 2022-07-24 cover.tex
minor changes

%% LyX 2.3.6.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{letter}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}
Dear Editor-In-Cheaf\medskip{}

we hereby submit our manuscript ``\emph{Multiset-trie data structure}''
with coauthors M. Akulich, I. Savnik, and R. ¦krekovski, for possible
publication in PLOS One.

\closing{Yours sincerely,\\
Matja¸ Krnc}
\end{document}