### 2022-08-13 tex/tex/s_conclusions.tex
paragraph added to s_conclusions.tex.

\section{Conclusions and future work} \label{c:conclusions}

%
% Experiments conclusions
One of the conclusions of studying the multiset-trie both theoretically and empirically is that our data structure is input sensitive. Input sensitivity implies a non-consistent performance on different input data. However, our argument that the performance can be optimized by pre-processing the input data is confirmed in  Experiment~\hyperref[ss:exp3]{4}. Pre-processing determines the optimal encoding for input data and ensures the best performance of the multiset-trie on particular input data. For example, in the case of storing words in the multiset-trie, the search queries can always be optimized based on the frequencies of letters in a specific language. We also see from Experiments~\hyperref[s:exp1]{1} and~\hyperref[s:exp2]{2} that the dependence of the multiset-trie performance on the density is not a linear function. Yet the function is continuous, and the point of inflection is unique on the whole domain, as shown in Experiment~\hyperref[s:exp3]{3}. This allows us to predict whether multiset-trie can be used for some particular application, serving a high performance.

%
% Mathematical analysis conclusions
% The mathematical analysis of the space complexity shows that multiset-trie requires only $O(|M|)$ space, which is the minimal possible space required by any data structure for storage of $|M|$ objects. As for the running time complexity of algorithms,
The basic tree functions such as \textsc{insert}, \textsc{search}, and \textsc{delete} all have a constant complexity once the multiset-trie is defined. The "getAll" multiset containment functions have worst-case running time complexity of $O(|\mathcal{M}|),$ where $|\mathcal{M}|$ is the cardinality of the multiset-trie data structure. The "existence" multiset containment functions have the worst-case running time complexity of $O(|\mathcal{M}| - |M|),$ where $|\mathcal{M}|$ is the cardinality of the multiset-trie and $|M|$ is the number of inserted multisets (nodes on leaf level).

The multiset-trie is an input-sensitive data structure because the size of multiset-trie $|\mathcal{M}|$ depends on the distribution of multisets in $M.$
Our mathematical model assumes that multisets $m$ in $M$ are distributed uniformly;  however, in real-world models, such an assumption is not generally true. 
%Specifically, the probability $P(m\in M)$ may vary dramatically and can even be equal to $0.$ For example, if words are mapped to multisets, then the sample space contains very large multisets. 
For example, many multisets will have zero probability of appearing in $M$ because a word that would correspond to such a multiset does not exist in the dictionary.%

The implementation of the multiset-trie is not optimized. The main reason for space-inefficiency is in implementing the links from a node to its children. A node at the $i$-th level of a tree has $n-i$ possible children where $n$ is the size of the alphabet. Our implementation uses an array data structure to link a node to its children where each element of an array includes a link to a sub-tree. A custom-implemented small and extendable hash table would significantly decrease the amount the space needed, especially since the number of children tends to decrease with the increasing levels of a tree. There are more additional optimizations possible. For example, the ...

% Future work notes
%
%The above results have opened even more interesting questions for future research. 
Further steps in our research will be to extend the functionality of the multiset-trie. We are interested in more flexible multiset containment queries where additional conditions constrain the sub and super-multisets. For example, the multiplicity of an element in a multiset can be bounded in operations getAllSubmsets and getAllSupermsets. Furthermore, the similarity search on multisets can be implemented by modifying the algorithms for searching the sub and super-multisets. 
%
The second line of research is to investigate the multiset-trie as a database index data structure. A disk-based index data structure allows storing and managing a huge amount of multisets. The mapping from a multiset-trie, i.e., a $n$-ary search tree, to a block-based index can be easily defined because of the regularity of multiset-trie. It will be interesting to compare the multiset-trie with other existing disk-based index data structures.