### 2023-02-21 tex/tex/s_data_structure.tex
reponses to some comments and updates of corresponding tex files.

\section{Multiset-trie data structure} \label{c:description}
%
Let $\Sigma$ be a set of distinct symbols that define an alphabet, and let 
$\sigma$ be the cardinality of $\Sigma.$ The \emph{multiset-trie} data structure 
stores multisets that are composed of symbols from the alphabet $\Sigma.$ It 
provides the basic tree data structure operations such as insert, delete, and 
search, together with multiset containment operations for searching sub-multisets
and super-multisets that will be discussed in the next section in greater detail.

A multiset ignores the ordering of its elements by definition, which allows us to 
define a bijective mapping $\phi:\Sigma \rightarrow I,$ where $I$ is the set of 
integers $\{ 1,2,3, \ldots, \sigma \}.$ In this way, we obtain indexing of 
elements from the alphabet $\Sigma,$ so we can work directly with integers 
rather than with specific symbols from $\Sigma.$

The multiset-trie is an $n$-ary tree-based data structure with the properties of 
the trie. A node in multiset-trie always has degree $n,$ i.e., $n$ children. Some of 
the children may be \emph{Null} (non-existing), but the number of \emph{Null} 
children can be at most $n-1.$ All the children of a node, including the \emph{Null} 
children, are labeled from left to right with labels $c_j,$ where $j\in \{ 0, 1, \ldots, n-1 \}.$ 
Every pair of child nodes $u$ and $v$ that share the same parent node have different labels.

%The multiset-trie is an $n$-ary tree based structure. A node can have a degree 
%from $1$ to $n,$ depending on the number of existing children. The edges from 
%parent node to children have labels $c_j,$ where $j\in \{ 0, 1, \ldots, n-1 \}.$ 
%The edges are labeled from left to right with the condition that there are no 
%two edges with the same label coming from a parent node.

Nodes that have equal height in a multiset-trie form a level. The height of a 
multiset-trie is always $\sigma+1$ if at least one multiset is in structure. 
The height of the root node (the first level) is defined to be 1.
%
Levels in multiset-trie are enumerated by their height, i.e., a level $L_i$ has 
height $i.$ The connection between level height in a multiset-trie and symbols 
from alphabet $\Sigma$ is defined as follows. A level $L_i,$ where 
$i\in\{ 1,2,\ldots, \sigma \}$ represents a symbol $s\in\Sigma,$ such that 
$\phi^{-1}(i) = s.$ The last level $L_{\sigma+1}$ does not represent any symbol 
and is named \emph{leaf level} ($LL$ for short).

Since every level, except $LL$ represents a symbol from $\Sigma,$ we can define 
a transition between nodes that are located at different levels in a multiset-trie. 
%
Consider two nodes $u,v$ in a multiset-trie at levels $L_i, L_{i+1}$ respectively, 
where $i\in\{1,2,\ldots,\sigma\}.$ Let a node $u$ be a parent node of a node $v$ 
and consequently a node $v$ be a child node of a node $u.$ Suppose that a child 
node $v$ is not \emph{Null} and has a label $c_j,$ where $j\in\{ 0,1,\ldots, n-1 \}.$ 
%
Then the \emph{path} $u\rightarrow v$ represents a symbol $s\in\Sigma$ with 
multiplicity $j,$ such that $\phi^{-1}(i) = s.$ 
%
Such a transition $u\rightarrow v$ is called a \emph{path of length} $1$ and is 
allowed if and only if a node $v$ is not \emph{Null} and $u$ is a parent node of 
a node $v.$ If a node $v$ has label $c_0,$ then the path $u\rightarrow v$ 
represents a symbol with the multiplicity $0$ respectively i.e. an empty symbol.

We define a \emph{complete path} to be the path of length $\sigma$ in a 
multiset-trie with the endpoints at the root node (the 1st level) and $LL$. Thus, 
a multiset $m$ is inserted into a multiset-trie if and only if there exists a 
complete path in a multiset-trie that corresponds to $m.$
%
Note that every complete path in a multiset-trie is unique. Therefore, the multisets 
that share a common prefix in a multiset-trie can have a common path of length at 
most $\sigma-1.$ The complete path that passes through nodes labeled by $c_0$ 
on all levels represents an empty multiset or an empty set.
%
Thus, any multiset $m$ that is composed of symbols from $\Sigma$ with maximum 
multiplicity not greater than $n-1$ can be represented by a complete path in a multiset-trie.

An example of a multiset-trie data structure with $\sigma = 2$ and 
$\Sigma = I = \{ 1,2 \}$ (i.e. the mapping $\phi$ is an identity mapping) is shown on Figure~\ref{fig:sketch}. 
On the figure, which stores elements of $\{\emptyset, \{ 1,1,2 \},$ $\{ 1,2,2 \},$ $\{ 2 \},$ $\{ 1,2 \},$ $\{ 2,2 \}\}$, the degree of a node is set to be $n=3,$ so the maximal multiplicity of an element in 
a multiset is $n-1=2.$

\begin{figure}[h!]
\centering
\begin{tikzpicture}[>=stealth',
level 1/.style={sibling distance = 2.5cm},
level 2/.style={sibling distance = 1cm},
level distance = 1.5cm]
\node (R) [node1] {\tiny Root}
	child { node [node1] {} 
		child { node [node1] {} edge from parent node [near end, left] {$c_0$}}
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		child { node [node1] {} edge from parent node [near end, right] {$c_2$}}
		edge from parent node [near end, left] {$c_0$}	
	}
	child { node [node1] {}
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		child { node [node1] {} edge from parent node [near end, right] {$c_2$}}
		edge from parent node [near end, left] {$c_1$}
	}
	child { node [node1] {} 
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		edge from parent node [near end, right] {$c_2$} 
};	
%----------------- Level labels --------------	
	\begin{scope}[every node/.style={right}]
		\path (R 				-| R-3-1) ++(5mm,0) node {$L_1$};
		\path (R-3 			-| R-3-1) ++(5mm,0) node {$L_2$};
		\path (R-3-1		-| R-3-1) ++(5mm,0) node {$LL$};
	\end{scope}			
\end{tikzpicture}
\caption{Example of multiset-trie structure containing multisets multisets $\emptyset, \{ 1,1,2 \},$ $\{ 1,2,2 \},$ $\{ 2 \},$ $\{ 1,2 \},$ $\{ 2,2 \}$. The \emph{Null} children are omitted.}
\label{fig:sketch}
\end{figure}

Let a pair $(L_i, c_j)$ represents a node with label $c_j$ at a level $L_i.$ 
The pair $(L_1,c_j)$ is equivalent to $(L_1,root),$ since the first level has 
the root node only. According to Figure~\ref{fig:sketch} we can extract 
inserted multisets as follows:

\begin{eqnarray*}
(L_1,root) \rightarrow (L_2,c_0) \rightarrow (LL,c_0) & = & \{ 1^0, 2^0 \} = \emptyset \\
(L_1,root) \rightarrow (L_2,c_0) \rightarrow (LL,c_1) & = & \{ 1^0, 2^1 \} = \{ 2 \} \\
(L_1,root) \rightarrow (L_2,c_0) \rightarrow (LL,c_2) & = & \{ 1^0, 2^2 \} = \{ 2,2 \} \\
(L_1,root) \rightarrow (L_2,c_1) \rightarrow (LL,c_1) & = & \{ 1^1, 2^1 \} = \{ 1,2 \} \\
(L_1,root) \rightarrow (L_2,c_1) \rightarrow (LL,c_2) & = & \{ 1^1, 2^2 \} = \{ 1,2,2 \} \\
(L_1,root) \rightarrow (L_2,c_2) \rightarrow (LL,c_1) & = & \{ 1^2, 2^1 \} = \{ 1,1,2 \} 
\end{eqnarray*}
where $e^k$ represents an element $e$ with multiplicity $k.$